Embedding Dimension of Polyhedral Losses

### 2020-01-28 colt-20/embed-dim.tex
Start proof of abstain on 5 outcomes not 2 embeddable.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Higher dimensions}
\bo{blurb about the purpose of this section and the layout.}
\jessie{Taking a crack at it:}
Our newfound understanding of the 1-dimensional embedding characterization leads us to realize that a large class of properties are not $1$-embeddable.
\cite{finocchiaro2019embedding} have shown a trivial upper bound that every property is $(n-1)$-embeddable, but it is an open question as to how tight we can get the bounds on embedding dimension.
Reducing embedding dimension significantly below $(n-1)$ can help provide a computational speedup in the optimization problem, which is a function of the surrogate loss dimension.
This is especially important as the number of possible outcomes $n$ grows large.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Higher dimensions}\label{sec:d-dim}
\bo{blurb about the purpose of this section and the layout.}
\jessie{Taking a crack at it:}
Our newfound understanding of the 1-dimensional embedding characterization leads us to realize that a large class of properties are not $1$-embeddable.
\cite{finocchiaro2019embedding} have shown a trivial upper bound that every property is $(n-1)$-embeddable, but it is an open question as to how tight we can get the bounds on embedding dimension.
Reducing embedding dimension significantly below $(n-1)$ can help provide a computational speedup in the optimization problem, which is a function of the surrogate loss dimension.
This is especially important as the number of possible outcomes $n$ grows large.

@@

\begin{proposition}
	$\gamma := \abstain{1/2}$ with $n=5$ is not $2$-embeddable.
\end{proposition}
\begin{proof}
  We use the results from Section~\ref{sec:d-dim} and show that the optimality condition is not met for the level set $\gamma_\bot$.
  First, we represent the cell $\gamma_\bot = \{p \in \simplex : Bp \geq \vec 0\} = \conv(\{(1/2, 1/2, 0,0,0), \ldots, (0,0,0,1/2, 1/2)\})$, where $B = \ones \ones^T - 2I_5 \in \reals^{5 \times 5}$ has entry $-1$ on the diagonal, and entry $1$ everywhere else.

  We make an argument on the permissible angles between normals $v_1, \ldots, v_5$ in $\reals^2$ and show there is no feasible solution to the QFP in Definition~\ref{def:qfp}.
  In particular, we argue that for any cone strictly $ < 180^\circ$, we can have no more than two vectors $v_i$ and $v_j$; otherwise we yield a contradiction.

  Without loss of generality, we let $v_1 = (0,1)$, and have $v_2$ with $\|v_2\|^2 \leq \|v_1\|^2$ and the angle $\theta_{12} < 180^\circ$ the angle between $v_1$ and $v_2$.

  We then have the feasible regions $F_y = \{x \in \reals^2 : \inprod{v_i} {x} \leq B_{iy} \; \forall i\}$, which means to this point, we have $F_1 = \{x \in \reals^2 : \inprod{v_1}{x} \leq -1 \text{ and } \inprod{v_2}{x} \leq 1 \}$, and $F_2 = \{x \in \reals^2 : \inprod{v_1}{x} \leq 1 \text{ and } \inprod{v_2}{x} \leq -1 \}$.
  By requirement of the existence of a set of witnesses for the distribution $(1/2, 1/2, 0,0,0)$, we must have some $x^{12}_1 \in F_1$ and $x^{12}_2 \in F_2$ so that $x^{12}_1 = -x^{12}_2$ in order for $\vec 0$ to be in the Minkowski sum.
  In particular, we know that, since $\theta_{12} < 180^\circ$, $x^{12}_1$ and $x^{12}_2$ are unique given $v_1$ and $v_2$.

  We then want to show that we cannot add a $v_3$ such that $\max(\theta_{12}, \theta_{13}, \theta_{23}) < 180^\circ$ without yielding a contradiction, which then gives our result, as any arrangement of $5$ normals in $\reals^2$ must have some cone that is less than $180^\circ$ containing three normals.

  In general, observe that $x^{12}_1$ and $x^{13}_1$ must both be on the line $\inprod{v_1}{x} = -1$. \jessie{More detail}  
  Therefore, our three cases proceed as follows: $(i.) \|x^{13}_1\| > \|x^{12}_1\|$, $(ii.) \|x^{13}_1\| = \|x^{12}_1\|$, and $(iii.) \|x^{12}_1\| > \|x^{13}_1\|$.
  As the labels on normals $v_2$ and $v_3$ (and therefore, the respective witnesses), it suffices to just show cases $(i.)$ and $(ii.)$.

  \begin{enumerate}
  \item [$(i.)$] First, suppose $\|x^{13}_1\| < \|x^{12}_1\|$.
  This in turn implies that $\|v_3\| > \|v_2\|$ since both witnesses are on the line $\inprod{v_1}{x^{1i}_1} = -1$.
  Consider that in order for $x^{12}_1 \in F_1$, we must have both $\inprod{v_2}{x^{12}_1} = 1$ and $\inprod{v_3}{x^{12}_1} \leq 1$.
    \begin{align*}
    \inprod{v_2}{x^{12}_1}} = 1 &\iff \|v_2\| \cos(\theta(v_2, x^{12}_1)) = \frac 1 {\|x^{12}_2\|}\\
    \inprod{v_3}{x^{12}_1}} \leq 1 &\iff \|v_3\| \cos(\theta(v_3, x^{12}_1)) \leq \frac 1 {\|x^{12}_2\|}\\
\implies \|v_3\| \cos(\theta(v_3, x^{12}_1)) &\leq \|v_2\| \cos(\theta(v_2, x^{12}_1))\\
\implies 1 < \frac{\|v_3\|}{\|v_2\|} &\leq \frac{\cos(\theta(v_2, x^{12}_1))}{\cos(\theta(v_3, x^{12}_1))}\\
\implies \theta(v_3, x^{12}_1) &> \theta(v_2, x^{12}_1)\\
\implies \theta_{13} < \theta_{12}
    \end{align*}
    This yields a contradiction, as we must have $\theta_{12} < \theta_{13}$ in order to have $\inprod{v_3}{x^{13}_1} = \inprod{v_2}{x^{12}_1} = 1$.
    Therefore, we conclude that we cannot have $\|x^{12}_1\| > \|x^{13}\|$.
  	\item [$(ii.)$] Suppose that we have $\|x^{12}_1\| = \|x^{13}_1\|$.
  	As both witnesses are on the same halfspace $\{x : \inprod{v_1}{x} = -1\}$, we must have $x^{12}_1 = x^{13}_1$.

  \end{enumerate}

%  We do this by arguing that in all three cases of the argmax, we either restrict $F_1$ so that $x_1 \not \in F_1$, restrict $F_2$ so that $x_2 \not \in F_2$, or have $x_3 \not \in F_3$.
%  We merge the latter two cases without loss of generality as we can relabel these normals pretty easily because of the structure of $B$.
%

%  
%  \begin{enumerate}
%  	\item [$\theta_{23} = \argmax$] 
%  	Since $\theta_{23} < 180^\circ$ and $v_1$ is in the middle of $v_2$ and $v_3$, we have that $x^{12}_1 \neq x^{13}_1$.
%  	Without loss of generality, suppose $\|v_2\|^2 \leq \|v_3\|^2$.
%  	Then we claim that $x^{13}_1 \not \in F_1$, since we have $x^{13}_1 = \{x : \inprod{v_1}{x} = -1 \text{ and } \inprod{v_3}{x} = 1\}$.
%  \end{enumerate}

\end{proof}
